%==============================================================================
\appendix
\section{Rendezvous Agent --- GLP Implementation}
\label{appendix:rv-agent}
%==============================================================================
The rendezvous agent processes cold-call messages on its input stream.  A $\mathit{reconnect}$ message from agent~$A$ includes~$A$'s public key, friend~$B$'s public key, a friendship proof, and a reply variable.  An $\mathit{available}$ message from~$B$ includes both public keys and a reply variable.

The agent verifies the friendship proof via \code{verify\_friendship} and observes each connecting agent's public address via \code{peer\_address} (both body system predicates).  It searches a pending list for a matching request from the other side.  If found, \code{complete\_rendezvous} constructs the reply values in its head: $A$ receives $\mathit{punch}(\mathit{Addr}_B)$ and $B$ receives $\mathit{punch}(\mathit{Addr}_A)$.  If the proof is invalid, the reply is bound to $\mathit{rejected}$.  If no match is found, the request is stored in the pending list.

The key GLP mechanism is the \emph{reply-variable chain}: each cold-call message contains a reply variable (a writer held by the sender).  As the rendezvous agent processes the message, the reply variable is threaded through writer$\to$reader handoffs.  When both sides are matched, the assignments propagate back through the chain to the original senders.

\begin{verbatim}
SignalReply ::= punch(Constant) ; rejected.
VerifyResult ::= ok ; invalid.

SignalMsg ::= reconnect(Constant, Constant, Constant, SignalReply?)
              %% reconnect(A_pk, B_pk, Proof, Reply)
            ; available(Constant, Constant, SignalReply?).
              %% available(B_pk, A_pk, Reply)

PendingEntry ::= needs(Constant, Constant, Constant, SignalReply?)
                 %% needs(A_pk, B_pk, AddrA, ReplyA?)
               ; ready(Constant, Constant, Constant, SignalReply?).
                 %% ready(B_pk, A_pk, AddrB, ReplyB?)
PendingList ::= [] ; [PendingEntry | PendingList].

procedure rv_agent(Stream(SignalMsg)?, PendingList?).

rv_agent([reconnect(A, B, Proof, ReplyA?)|In], Pending) :-
    ground(A?), ground(B?) |
    verify_friendship(Proof?, A?, B?, Status),
    peer_address(A?, AddrA),
    handle_verify(Status?, A?, B?, AddrA?, ReplyA, Pending?, Pending1),
    rv_agent(In?, Pending1?).

rv_agent([available(B, A, ReplyB?)|In], Pending) :-
    ground(A?), ground(B?) |
    peer_address(B?, AddrB),
    find_needs(A?, B?, AddrB?, ReplyB, Pending?, Pending1),
    rv_agent(In?, Pending1?).

rv_agent([], _).

procedure handle_verify(VerifyResult?, Constant?, Constant?, Constant?,
                        SignalReply, PendingList?, PendingList).
handle_verify(ok, A, B, AddrA, ReplyA?, Pending, Pending1?) :-
    find_ready(A?, B?, AddrA?, ReplyA, Pending?, Pending1).
handle_verify(invalid, _, _, _, rejected, Pending, Pending?).

procedure find_ready(Constant?, Constant?, Constant?, SignalReply,
                     PendingList?, PendingList).
find_ready(A, B, AddrA, ReplyA?, [ready(B2, A2, AddrB, ReplyB?)|Ps], Ps?) :-
    A? =?= A2?, B? =?= B2? |
    complete_rendezvous(AddrA?, AddrB?, ReplyA, ReplyB).
find_ready(A, B, AddrA, ReplyA?, [P|Ps], [P?|Ps1?]) :-
    otherwise |
    find_ready(A?, B?, AddrA?, ReplyA, Ps?, Ps1).
find_ready(A, B, AddrA, ReplyA?, [], [needs(A?, B?, AddrA?, ReplyA)]).

procedure find_needs(Constant?, Constant?, Constant?, SignalReply,
                     PendingList?, PendingList).
find_needs(A, B, AddrB, ReplyB?, [needs(A2, B2, AddrA, ReplyA?)|Ps], Ps?) :-
    A? =?= A2?, B? =?= B2? |
    complete_rendezvous(AddrA?, AddrB?, ReplyA, ReplyB).
find_needs(A, B, AddrB, ReplyB?, [P|Ps], [P?|Ps1?]) :-
    otherwise |
    find_needs(A?, B?, AddrB?, ReplyB, Ps?, Ps1).
find_needs(A, B, AddrB, ReplyB?, [], [ready(B?, A?, AddrB?, ReplyB)]).

procedure complete_rendezvous(Constant?, Constant?, SignalReply, SignalReply).
complete_rendezvous(AddrA, AddrB, punch(AddrB?), punch(AddrA?)).
\end{verbatim}

The system predicates \code{peer\_address/2}, \code{verify\_friendship/4}, and \code{punch\_udp/1} are described in Section~\ref{sec:rv-predicates}.

%==============================================================================
\section{Rendezvous Client --- GLP Implementation}
\label{appendix:rv-client}
%==============================================================================
The client side provides building blocks for agents that need to use a rendezvous agent.  \code{make\_reconnect} constructs a $\mathit{reconnect}$ cold-call message with a fresh reply variable embedded, returning the message and the paired reply reader.  \code{make\_available} does the same for $\mathit{available}$ messages.  \code{handle\_reply} dispatches on the reply: $\mathit{punch}(\mathit{Addr})$ triggers \code{punch\_udp}; $\mathit{rejected}$ requires no networking action.

\begin{verbatim}
SignalReply ::= punch(Constant) ; rejected.

SignalMsg ::= reconnect(Constant, Constant, Constant, SignalReply?)
              %% reconnect(A_pk, B_pk, Proof, Reply)
            ; available(Constant, Constant, SignalReply?).
              %% available(B_pk, A_pk, Reply)

procedure make_reconnect(Constant?, Constant?, Constant?,
                         SignalMsg, SignalReply).
make_reconnect(A, B, Proof,
               reconnect(A?, B?, Proof?, Reply),
               Reply?).

procedure make_available(Constant?, Constant?, SignalMsg, SignalReply).
make_available(B, A,
               available(B?, A?, Reply),
               Reply?).

procedure handle_reply(SignalReply?).
handle_reply(punch(Addr)) :- ground(Addr?) | punch_udp(Addr?).
handle_reply(rejected).
\end{verbatim}

An end-to-end test wires the client and server together on a shared message stream:
\begin{verbatim}
procedure test_full_scenario(SignalReply, SignalReply).
test_full_scenario(ReplyA?, ReplyB?) :-
    make_reconnect(alice, bob, proof, MsgA, ReplyA),
    make_available(bob, alice, MsgB, ReplyB),
    rv_agent([MsgA?, MsgB?], []).
\end{verbatim}
After execution, $\mathit{ReplyA}$ is bound to $\mathit{punch}(\mathit{stub\_addr\_bob})$ and $\mathit{ReplyB}$ to $\mathit{punch}(\mathit{stub\_addr\_alice})$.

%==============================================================================
\section{Friends-Based Rendezvous --- GLP Implementation}
\label{appendix:fmrv}
%==============================================================================
Friends-based rendezvous is implemented as an extension of the social graph agent.  The friends list is extended to include each friend's current IP address: entries have the form $\mathit{fe}(\mathit{Name}, \mathit{OutputStream}, \mathit{Addr})$.  Agents notify friends of address changes via $\mathit{addr\_update}$ messages.

When agent~$A$ asks friend~$B$ to mediate reconnection to~$C$, the procedure \code{lookup\_and\_send} performs a fused lookup-and-send: in a single traversal of the friends list, it locates a friend's entry, prepends a $\mathit{punch\_to}$ message to that friend's output stream, and returns the friend's address.  The $\mathit{punch\_to}$ argument carries an $\mathit{AddrResult}$: $\mathit{found}(\mathit{Addr})$ if the peer was located, or $\mathit{not\_found}$ if not.  The recipient dispatches accordingly --- \code{punch\_udp} is never called with an invalid address.

The $\mathit{mediate}$ clause calls \code{lookup\_and\_send} twice with cross-references: the first call sends $A$ a $\mathit{punch\_to}(\mathit{AddrC}?)$ where $\mathit{AddrC}$ is not yet bound; the second call binds $\mathit{AddrC}$ when it finds~$C$'s entry.  GLP's suspension mechanism fills in the unbound reader once the writer is bound --- the standard futures/promises mechanism.

\begin{verbatim}
AddrResult ::= found(Constant) ; not_found.

MsgOp ::= mediate(Constant)        %% mediate(C_pk)
        ; punch_to(AddrResult)     %% punch_to(found(Addr)) or punch_to(not_found)
        ; addr_update(Constant).   %% addr_update(NewAddr)

Msg ::= msg(Constant, Constant, MsgOp).
MsgStream ::= [] ; [Msg | MsgStream].
FriendStream ::= [] ; [Msg | FriendStream].
FriendEntry ::= fe(Constant, FriendStream?, Constant).
FriendList ::= [] ; [FriendEntry | FriendList].

procedure fmrv(Constant?, MsgStream?, FriendList?).

fmrv(Id, [msg(From, To, addr_update(Addr))|In], Fs) :-
    ground(Id?), Id? =?= To? |
    update_addr(From?, Addr?, Fs?, Fs1),
    fmrv(Id?, In?, Fs1?).

fmrv(Id, [msg(From, To, mediate(C))|In], Fs) :-
    ground(Id?), ground(From?), ground(C?), Id? =?= To? |
    lookup_and_send(From?, Id?, AddrC?, AddrFrom, Fs?, Fs1),
    lookup_and_send(C?, Id?, AddrFrom?, AddrC, Fs1?, Fs2),
    fmrv(Id?, In?, Fs2?).

fmrv(Id, [msg(_, To, punch_to(found(Addr)))|In], Fs) :-
    ground(Id?), Id? =?= To? |
    punch_udp(Addr?),
    fmrv(Id?, In?, Fs?).

fmrv(Id, [msg(_, To, punch_to(not_found))|In], Fs) :-
    ground(Id?), Id? =?= To? |
    fmrv(Id?, In?, Fs?).

fmrv(_, [], _).

procedure lookup_and_send(Constant?, Constant?, AddrResult?, AddrResult,
                          FriendList?, FriendList).
lookup_and_send(Key, Id, OtherAddr, found(Addr?),
    [fe(K, [msg(Id?, K?, punch_to(OtherAddr?))|Out1?], Addr)|Rest],
    [fe(K?, Out1, Addr?)|Rest?]) :-
    Key? =?= K?, ground(Addr?) | true.
lookup_and_send(Key, Id, OtherAddr, KeyAddr?,
    [F|Rest], [F?|Rest1?]) :-
    otherwise |
    lookup_and_send(Key?, Id?, OtherAddr?, KeyAddr, Rest?, Rest1).
lookup_and_send(_, _, _, not_found, [], []).

procedure update_addr(Constant?, Constant?, FriendList?, FriendList).
update_addr(Key, NewAddr, [fe(K, Out?, _)|Fs],
            [fe(K?, Out, NewAddr?)|Fs?]) :-
    Key? =?= K? | true.
update_addr(Key, NewAddr, [F|Fs], [F?|Fs1?]) :-
    otherwise | update_addr(Key?, NewAddr?, Fs?, Fs1).
update_addr(_, _, [], []).
\end{verbatim}

%==============================================================================
\section{IP Cold-Call Invite Redemption}
\label{appendix:ip-cold-call-invites}
%==============================================================================
An IP cold-call peer link is a signed bearer invitation created by agent~$A$ for an as-yet-unknown holder~$B$.  The link is base64url-encoded and is redeemed through a rendezvous server.

\begin{verbatim}
Invite ::= invite(
    A_pk,
    A_nickname,
    RvServer,        %% RV address and public key
    Nonce256,
    ExpiresAt,
    SignatureA
).
\end{verbatim}

\noindent \code{Nonce256} is 256 uniformly random bits.  \code{ExpiresAt} is one hour after creation. \code{RvServer} contains a rendezvous server's public address and key. \code{SignatureA} is an Ed25519 signature by~$A$ over all preceding fields. The nonce is usable exactly once: $A$ records the invite as unused with the rendezvous agent when the link is generated, and the rendezvous agent marks it consumed atomically when accepting the first-contact payload.  Possession of the link authorizes a first-contact attempt; friendship is established only when $A$ accepts the resulting cold call from $B$.

The link is base64-encoded for compact URI transport, allowing the recipient to open it as a deep-link.  The protocol relies on the out-of-band channel to deliver the link to the intended recipient, and relies on the 256-bit nonce, one-hour expiration, and $A$'s signature to prevent guessing, replay, and tampering.

Both $A$ and any holder of the link derive an opaque invite identifier and proof key:
\[
\begin{aligned}
\mathit{InviteId}  &= H(\text{``glp invite id''} \mid \mathit{Nonce256}),\\
\mathit{InviteKey} &= \mathit{HKDF}(\mathit{Nonce256},
    \text{``glp invite proof''} \mid A_{\mathit{pk}} \mid
    \mathit{RvServer} \mid \mathit{ExpiresAt}).
\end{aligned}
\]
The raw nonce is not sent over the IP network.  Network messages use \code{InviteId} plus MACs under \code{InviteKey}, so passive observers and off-path attackers cannot learn or guess the nonce from redemption traffic.

\subsection{Rendezvous-Gated Redemption}
\label{appendix:rv-invite-redemption}

\begin{enumerate}
\item \textbf{Generation and registration.}  $A$ picks a fresh 256-bit secret, computes the invite ID and proof key from it, and assembles the link (public key, nickname, the chosen rendezvous server~$S$'s address and public key, secret, expiration).  $A$ signs the link with its private key and records the invite locally as unused until the expiration.  Over $A$'s existing authenticated connection to~$S$, $A$ registers the invite by sending $A$'s public key, the invite ID, the proof key, and the expiration.  $S$ stores these together with an ``unused'' flag.

\item \textbf{Link delivery.}  $A$ sends the base64url-encoded link to~$B$ out-of-band.

\item \textbf{Cold-call claim.}  $B$ opens the link.  \func{consumePeerLink(uri)} decodes it, verifies $A$'s signature, checks that it is still valid, registers $A$'s public key and nickname locally, and recomputes the invite ID and proof key from the secret.  $B$ opens an encrypted, authenticated connection to~$S$ (using $S$'s public key from the link) and submits a cold-call claim: the invite ID, $B$'s public key, $B$'s nickname, and a short proof computed from the proof key covering $A$'s public key, $B$'s public key, the invite ID, and a fresh nonce of $B$'s own.  $B$ then sends a standard rendezvous \code{available} message naming itself and~$A$.

\item \textbf{Claim validation and forwarding.}  $S$ checks four conditions: the invite ID is registered for~$A$; the expiration is still in the future; the invite is still marked unused; and the proof verifies against the proof key stored in step~1.  On any failure, $S$ rejects the claim.  On success, $S$ atomically flips the invite to ``used'' and forwards $B$'s public key and nickname to~$A$ over $A$'s existing connection to~$S$.

\item \textbf{Acceptance.}  $A$ presents $B$'s public key and nickname to the application layer as a friendship request.  If $A$ accepts the request (which is the friendship decision), $A$ sends~$S$ a standard rendezvous \code{reconnect} message naming~$B$.  In place of the friendship attestation that the reconnect message normally carries, the proof field references the invite ID that $A$ pre-registered with~$S$ in step~1; $S$ matches it against its invite state to authorize the request.

\item \textbf{Signaling.}  $S$ now holds $A$'s \code{reconnect} and $B$'s \code{available} for the same pair of public keys, and the standard rendezvous protocol (Section~\ref{sec:rv-protocol}) takes over with~$S$ as the signaling agent.  $S$ observes both public addresses from the connections it holds and sends each side the other's address.

\item \textbf{UDP path and Noise handshake.}  $A$ and~$B$ each send UDP packets toward the address they received.  Once the UDP path is open, the side designated as initiator (by a deterministic rule based on the public keys) opens a UDX stream, and both sides run the Noise~XX handshake using their Ed25519 keys (Section~\ref{sec:ip-connection}).  The rendezvous server sees signaling metadata and observed addresses, never the encrypted payload.

\item \textbf{Cold-call delivery.}  Only after the Noise-authenticated UDX stream is established does $B$ send the madGLP cold-call payload, which the networking layer delivers to $A$'s index-0 serializer.  The friendship between $A$ and $B$ was established by $A$'s acceptance in step~5.
\end{enumerate}
