%==============================================================================
\section{BLE Medium}
\label{sec:ble}
%==============================================================================
This section specifies how the networking API (Section~\ref{sec:api}) is realized over Bluetooth Low Energy.
%------------------------------------------------------------------------------
\subsection{BLE Identity and Advertising}
\label{sec:ble-identity}
%------------------------------------------------------------------------------
Each agent advertises a BLE GATT service whose 128-bit (16-byte) UUID has two parts: a fixed 8-byte \emph{Grassroots prefix} common to all Grassroots agents, and an 8-byte \emph{agent suffix} derived from the agent's public key (e.g., the first 8 bytes of its SHA-256 hash). The prefix enables general discovery of any nearby Grassroots agent; the suffix enables recognition of a specific agent before connection, by anyone who knows that agent's public key.
%------------------------------------------------------------------------------
\subsection{BLE Discovery}
\label{sec:ble-discovery}
%------------------------------------------------------------------------------
The BLE layer maintains a continuous scan for devices whose service UUID matches the Grassroots prefix. A scan match is only a discovery hint: the advertisement carries neither the agent's public key nor its nickname, so it does not by itself constitute peer discovery.
Upon successful BLE connection, an ANNOUNCE packet is exchanged containing the agent's full 32-byte public key and a human-readable nickname. \func{onPeerDiscovered(cb)} fires upon first receipt of a peer's signed ANNOUNCE --- not on a scan match --- once per identity (Section~\ref{sec:discovery}); \func{getPeers()} reports the currently reachable peers among those so identified. No Bluetooth bonding or pairing confirmation is required; application-level encryption (using the Ed25519 keys) is used instead.

A BLE encounter with an unknown nearby agent is a local cold-call opportunity. Whether the agent accepts new BLE connections or not is governed by the environmental trust levels in Section~\ref{sec:cold-call-trust}.
%------------------------------------------------------------------------------
\subsection{BLE Communication}
\label{sec:ble-comm}
%------------------------------------------------------------------------------
\func{send(pubKey, payload)} transmits the byte payload over the established BLE GATT connection. The BLE networking layer handles fragmentation and reassembly for payloads exceeding the BLE MTU. \func{onReceive(cb)} delivers reassembled payloads to the runtime.