%==============================================================================
\section{IP Medium}
\label{sec:ip}
%==============================================================================
This section specifies how the networking API (Section~\ref{sec:api}) is realized over the Internet.  Communication uses UDP with UDX (reliable multiplexed streams over UDP) for data transport.  NAT traversal is handled by the rendezvous protocol (Section~\ref{sec:rendezvous}).

%------------------------------------------------------------------------------
\subsection{IP Connection}
\label{sec:ip-connection}
%------------------------------------------------------------------------------
Once a UDP path is established (either directly or via hole-punching), agents open a UDX stream over it. UDX provides reliable, ordered, congestion-controlled delivery over UDP, satisfying the networking assumptions (Section~\ref{sec:assumptions}).

Upon establishing the UDX stream, the peers perform a mutual authentication handshake using their Ed25519 keys via the Noise protocol XX pattern (Curve25519 for key exchange, ChaCha20-Poly1305 for encryption, SHA-256 for hashing).  The XX pattern transmits each peer's static public key during the handshake, bound to a proof of possession, and derives forward-secret session keys.  Each peer then verifies the delivered key against an expected key: the friend's known public key from the social graph for reconnection, or---for an IP cold-call---the inviter's public key from the invite (for the link holder) and the redeemer's public key from the RV-forwarded claim (for the inviter).  A mismatch aborts the connection.  The X25519 static key carried by Noise is the birational image of the agent's Ed25519 identity key (the standard Ed25519$\to$Curve25519 map), so the expected static is computable from the peer's known Ed25519 public key alone; no separate key distribution is required.  \dan{Implemented via libsodium's \code{crypto\_sign\_ed25519\_pk\_to\_curve25519} / \code{sk\_to\_curve25519}: each side recomputes the expected static from the verified sender Ed25519 key and aborts the handshake on mismatch (\code{NoiseSessionManager.\_verifyRemoteStatic}).  Defence-in-depth: every ANNOUNCE and handshake packet is also Ed25519-signed and verified against its claimed sender, and the Noise session is keyed by that verified identity.}

\func{onPeerConnected(cb)} fires when the UDX stream is established and authenticated. \func{onPeerDisconnected(cb)} fires when the stream drops.

%------------------------------------------------------------------------------
\subsection{IP Cold-Call}
\label{sec:ip-cold-call}
%------------------------------------------------------------------------------

Over IP, a cold call is bootstrapped by a deep-link sent out-of-band from inviting agent~$A$ to some holder~$B$.  The link is a signed bearer invitation: at creation time $A$ need not know~$B$'s public key, and possession of a valid unused link authorizes exactly one first-contact attempt.

\func{generatePeerLink()} produces the link and registers the invite with a chosen rendezvous (RV) server. \func{consumePeerLink(uri)} verifies $A$'s signature, registers~$A$'s identity, and submits a cold-call claim to the RV server named in the link.  The RV server validates the claim and forwards $B$'s identity to~$A$; if $A$ accepts the friendship request, the standard rendezvous protocol (Section~\ref{sec:rv-protocol}) ensues with the RV server as the signaling agent.  Once the UDP path is open, the peers run UDX and Noise~XX (Section~\ref{sec:ip-connection}), after which the madGLP cold-call payload is delivered to $A$'s index-0 serializer.  Full invite mechanics are specified in Appendix~\ref{appendix:ip-cold-call-invites}.

%------------------------------------------------------------------------------
\subsection{IP Communication}
\label{sec:ip-comm}
%------------------------------------------------------------------------------
\func{send(pubKey, payload)} transmits the byte payload over the authenticated UDX stream. Each message is length-prefixed (4-byte big-endian length followed by payload bytes) to provide message framing. \func{onReceive(cb)} delivers complete, deframed messages to the runtime.
